---
title: "Project Proposal"
author: 'Isabel Siebrecht, Franklin Feng, & Kaylin Hernandez'
date: "October 11th, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: 'due October 11, 2021 by 11:59 PM '
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Load Data
```{r load-data, message = FALSE}
hdi1 <- read.csv("../data/HDI_Table1.csv")
hdi3 <- read.csv("../data/HDI_Table3.csv")
hdi4 <- read.csv("../data/HDI_Table4.csv")
hdi5 <- read.csv("../data/HDI_Table5.csv")
hdi7 <- read.csv("../data/HDI_Table7.csv")
hdi8 <- read.csv("../data/HDI_Table8.csv")
hdi9 <- read.csv("../data/HDI_Table9.csv")
hdi10 <- read.csv("../data/HDI_Table10.csv")
adolescent_births <- read.csv("../data/Adolescent_Births.csv")
contraceptive_prevalence <- read.csv("../data/Contraceptive_Prevalence.csv")
```

# Introduction and Data, including Research Questions

## Topic: Teen Pregnancy
## Research question 
How does a country’s quality of life influence the birth rates for teenagers 
aged 15 to 19 years old? In other words, how does a country’s human and 
economic resources affect the rate of premature births? We will define how we 
are choosing to define quality of life later in the proposal. 

Our group is very interested in looking into teenage pregnancies because we 
feel that adolescent pregnancies are one of the most significant factors that 
contribute to the status of development and prosperity a country has. Often 
defined as a social problem, many studies like this one have found links 
between teenage pregnancies and “economic well-being,” which could indirectly 
impact the labor market and economy of a country in the long term if the 
proportion of adolescent mothers is large enough. 

We are trying to see the correlation between assumed “quality of life” factors 
and teenage pregnancies for girls aged 15-19. To do this, we pulled data on 
adolescent birth rate (WHO dataset), which was collected through a variety of 
national and international surveys and country reports. We also pulled data on 
female educational attainment, GNI, life expectancy at birth, HDI, Labor force 
participation rate, and gender inequality index (Human Development Index), 
which is from the 2020 Human Development Report provided by the United Nations. 
We chose to look at these factors based on common risk factors that influence 
teenage pregnancies, which broadly relate to the individual, family, and 
community (Risk and Protective Factors). 

## Variables:
Adolescent birth rate (proportion out of population)
Country
Year / (3.5) Mean years of schooling
Prevalence of contraceptives / (4.5) Maternal Mortality Rate
Educational attainment (female)
GNI (Gross National Income)
Life expectancy at birth
HDI 
Labour Force Participation Rate
Gender inequality index 


# Glimpse

(Please use `glimpse` for your data, uploaded into the data folder, here.)
```{r glimpse datasets}
glimpse(hdi1)
glimpse(hdi3)
glimpse(hdi4)
glimpse(hdi5)
glimpse(hdi7)
glimpse(hdi8)
glimpse(hdi9)
glimpse(hdi10)
glimpse(adolescent_births)
glimpse(contraceptive_prevalence)
```

# Data Analysis Plan

(Specify the outcome (response, Y) and predictor (explanatory, X) variables you will use to answer your question, as well as the comparison groups you will use, if applicable. You may include very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. Note the statistical method(s) that you believe will be useful in answering your question(s). 
What results from these specific statistical methods are needed to support your hypothesized answer?)

Since the repercussions of teenage pregnancy can spread into many aspects of 
life, we have chosen a few different explanatory variables to compare how these 
differences may look in different settings. For the first part of our project, 
our explanatory variables will be year and country, while the response variable 
will be adolescent birth rates. This will give us a good idea of the 
prevalence/proportion of birth rates around the world. For the second part of 
the project, we will try to find factors that could correlate to higher birth 
rates (birth rates as the response variable), so the explanatory  variables 
would be adolescent birth rate, prevalence of contraceptives, educational 
attainment (female v. male), GDP, life expectancy at birth, HDI, women 
inequality index and women in the workforce. We chose to look at the prevalence 
of contraceptives because we are assuming a country that has better access to 
contraceptives will have lower rates of teen pregnancies. We chose GDP, life 
expectancy and level of education attainment because we wanted to measure 
quality of life in countries and a country's economy and level of education are 
both great indicators of the quality of life for the citizens. HDI, human 
development index, is another way of measuring quality of life that looks at 
the index of life expectancy, education, and per capita income. The women 
inequality index and women in the workforce rate are both good ways to measure 
the effect of adolescent pregnancy on a country because they focus on the 
wellbeing of women, the people most affected by teen pregnancy. 

As a side note, we are worried that the contraceptive prevalence data will be 
tricky to use because it is from a different set and the data isn’t as clean as 
the other data we found so we have chosen a backup variable incase once we start 
to wrangle our data it looks to be too complicated. We have chosen maternal 
mortality rate because we hypothesize that with women starting their 
childbearing at a younger age, they will have more medical complications and 
will be more likely to have more children, thus increasing the risk of dying 
from childbirth. 
We want to make ten graphs. The first nine will all have the same x variables 
but the y variables will differ so that we can compare. The tenth will have the 
prevalence of contraceptives as x and birth rate as y. For the level of 
education it will be important to make a distinction between female and male 
because teen pregnancies tend to hold back women’s education more than men’s. 
We can do this using a filter on color by gender. 
For analysing our data we want to look at conditional probability to test for 
correlation. We want to see if there is a correlation between GDP, level of 
education, prevalence of contraceptives and life expectancy and test each one 
against adolescent birth rate. To do this, we can utilize linear regression 
models and check the strength of the correlation with r – we would also have 
to find the best line of fit, working with r2 as well. 
We also want to check our assumptions about possible correlators, so we plan 
to use hypothesis testing, we will use a two sided t-test, and analyze  p-values 
for our data sets. The p-value is a conditional probability that tells us how 
likely it is to get our data again (or a more extreme version) given our null 
hypothesis is true. If our reasoning is correct, we expect there to be a 
correlation between the four variables above and adolescent birth rate. This 
means that we will need to reject the null hypothesis, with statistical evidence
that there is no correlation between these variables.

