---
title: "Project Proposal"
author: 'Isabel Siebrecht, Franklin Feng, & Kaylin Hernandez'
date: "October 11th, 2021"
output:
  html_document:
    df_print: paged
  pdf_document: default
subtitle: 'due October 11, 2021 by 11:59 PM. Resubmitted October 18, 2021'
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
``` 

# Load Data
```{r load-data, message = FALSE}
hdi1 <- read.csv("../data/HDI_Table1.csv")
hdi4 <- read.csv("../data/HDI_Table4.csv")
hdi5 <- read.csv("../data/HDI_Table5.csv")
hdi8 <- read.csv("../data/HDI_Table8.csv")
hdi10 <- read.csv("../data/HDI_Table10.csv")
adolescent_births <- read.csv("../data/Adolescent_Births.csv")
contraceptive_prevalence <- read.csv("../data/Contraceptive_Prevalence.csv")
```


# Introduction and Data, including Research Questions

## Topic: Teen Pregnancy

## Research question 
How does a country’s quality of life influence the birth rates for teenagers 
aged 15 to 19 years old? In other words, how does a country’s human and economic
resources affect the rate of premature births? We will define how we are 
choosing to define quality of life later in the proposal. 

Our group is very interested in looking into teenage pregnancies because we feel
that adolescent pregnancies are one of the most significant factors that 
contribute to the status of development and prosperity a country has. Often 
defined as a social problem, many studies have found links between teenage 
pregnancies and “economic well-being,” which could indirectly impact the labor 
market and economy of a country in the long term if the proportion of adolescent
mothers is large enough. 

We are trying to see the correlation between assumed “quality of life” factors 
and teenage pregnancies for girls aged 15-19. To do this, we pulled data on 
adolescent birth rate (WHO dataset), which was collected through a variety of 
national and international surveys and country reports. We also pulled data on 
female educational attainment, GDP, life expectancy at birth, HDI, labour force 
participation rate, and gender inequality index (Human Development Index), which
is from the 2020 Human Development Report provided by the United Nations. We 
chose to look at these factors based on common risk factors that influence 
teenage pregnancies, which broadly relate to the individual, family, and 
community (Risk and Protective Factors).

## Variables:
Adolescent birth rate (proportion out of population)

Country

Year

Prevalence of contraceptives / (4.5) Mortality Rates, By Infants and Mothers

Educational Attainment (Female)

GDP (Gross Domestic Product)

Life Expectancy at Birth

HDI (Human Development Index)

Labour Force Participation Rate

Gender Inequality Index


# Glimpse

(Please use `glimpse` for your data, uploaded into the data folder, here.)
```{r glimpse datasets}
glimpse(hdi1)
glimpse(hdi4)
glimpse(hdi5)
glimpse(hdi8)
glimpse(hdi10)
glimpse(adolescent_births)
glimpse(contraceptive_prevalence)
```

Note: The datasets in the glimpse setting seem to be a little messy. However, 
with our feedback from Jackie Du we will use some data-wrangling and dplyr to
make this a more manageable and clean output. The commented out code is where we 
are looking to begin to clean up each dataset.


# Data Analysis Plan

Since the repercussions of teenage pregnancy can spread into many aspects of 
life, we have chosen a few different explanatory variables to compare how these 
differences may look in different settings. For the first part of our project, 
our explanatory variables will be year and country, while the response variable 
will be adolescent birth rates. This will give us a good idea of the 
prevalence/proportion of birth rates around the world. 

For the second part of the project, we will try to find factors that could 
correlate to higher birth rates (birth rates as the response variable), so the 
explanatory variables would be adolescent birth rate, prevalence of 
contraceptives, educational attainment (female v. male), GDP, life expectancy at
birth, HDI, gender inequality index and labour force participation rate 
(female v. male). We chose to look at the prevalence of contraceptives because 
we are assuming a country that has better access to contraceptives will have 
lower rates of teen pregnancies. We chose GDP, life expectancy and level of 
education attainment because we wanted to measure quality of life in countries 
and a country's economy and level of education are both great indicators of the 
quality of life for the citizens. HDI, the Human Development Index, is another 
way of measuring quality of life that looks at the index of life expectancy, 
education, and per capita income. The gender inequality index and labour force 
participation rate are both good ways to measure the effect of adolescent 
pregnancy on a country because they focus on the wellbeing of women, the people 
most affected by teen pregnancy. 

As a side note, we are worried that the 
contraceptive prevalence data will be tricky to use because it is from a 
different set and the data isn’t as clean as the other data we found so we have 
chosen a backup variable incase once we start to wrangle our data it looks to be
too complicated. We have chosen mortality rates among infants and mothers 
because we hypothesize that with women starting their childbaring at a younger 
age, they will have more medical complications and will be more likely to have 
more children, thus increasing the risk of dying from childbirth. The mortality 
rate among infants is another added variable that we can explore to observe the 
safety and resources of a country for the newborn child.

We want to make ten visualizations in total. The first nine will all have the 
same x variables but the y variables will differ so that we can compare. The 
tenth will have the prevalence of contraceptives as x and birth rate as y. For 
the level of education it will be important to make a distinction between female
and male because teen pregnancies tend to hold back women’s education more than 
men’s. We can do this using a filter on color by gender. For analysing our data 
we want to look at conditional probability to test for correlation. We want to 
see if there is a correlation between GDP, level of education, prevalence of 
contraceptives and life expectancy and test each one against adolescent birth 
rate. To do this, we can utilize linear regression models and check the strength
of the correlation with r – we would also have to find the best line of fit, 
working with R2 as well. We also want to check our assumptions about possible 
correlators, so we plan to use hypothesis testing, we will use a two sided 
t-test, and analyze p-values for our data sets. The p-value is a conditional 
probability that tells us how likely it is to get our data again 
(or a more extreme version) given our null hypothesis is true. If our reasoning 
is correct, we expect there to be a correlation between the four variables above
and adolescent birth rate. This means that we will need to reject the null 
hypothesis, with statistical evidence that there is no correlation between these
variables.