---
title: "Project Proposal"
author: 'INSERT TEAM NAME: Isabel, Franklin, & Kaylin'
date: "October 11th, 2021"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: 'due October 11, 2021 by 11:59 PM '
---

# Load Packages

```{r load-packages, message = FALSE, warning = FALSE}
library(tidyverse)
``` 

# Load Data
```{r load-data, message = F}
load(data, hdi1)
load(data, hdi4)
load(data, hdi5)
load(data, contraceptives)
load(data, adolescent_births)
```

# Introduction and Data, including Research Questions

## Topic: Teen Pregnancy
## Research question 
How does a country’s quality of life influence the birth rates for teenagers 
aged 15 to 19 years old? In other words, how does a country’s human and 
economic resources affect the rate of premature births? We will define how we 
are choosing to define quality of life later in the proposal. 

Our group is very interested in looking into teenage pregnancies because we 
feel that adolescent pregnancies are one of the most significant factors that 
contribute to the status of development and prosperity a country has. Often 
defined as a social problem, many studies like this one have found links 
between teenage pregnancies and “economic well-being,” which could indirectly 
impact the labor market and economy of a country in the long term if the 
proportion of adolescent mothers is large enough. 

We are trying to see the correlation between assumed “quality of life” factors 
and teenage pregnancies for girls aged 15-19. To do this, we pulled data on 
adolescent birth rate (WHO dataset), which was collected through a variety of 
national and international surveys and country reports. We also pulled data on 
female educational attainment, GNI, life expectancy at birth, HDI, Labor force 
participation rate, and gender inequality index (Human Development Index), 
which is from the 2020 Human Development Report provided by the United Nations. 
We chose to look at these factors based on common risk factors that influence 
teenage pregnancies, which broadly relate to the individual, family, and 
community (Risk and Protective Factors). 

## Variables:
Adolescent birth rate (proportion out of population)
Country
Year / (3.5) Mean years of schooling
Prevalence of contraceptives / (4.5) Maternal Mortality Rate
Educational attainment (female)
GNI (Gross National Income)
Life expectancy at birth
HDI 
Labour Force Participation Rate
Gender inequality index 


# Glimpse

(Please use `glimpse` for your data, uploaded into the data folder, here.)
```{r glimpse datasets}
glimpse(hdi1)
glimpse(hdi4)
glimpse(hdi5)
glimpse(contraceptives)
glimpse(adolescent_births)
```

# Data Analysis Plan

(Specify the outcome (response, Y) and predictor (explanatory, X) variables you will use to answer your question, as well as the comparison groups you will use, if applicable. You may include very preliminary exploratory data analysis, including some summary statistics and visualizations, along with some explanation on how they help you learn more about your data. Note the statistical method(s) that you believe will be useful in answering your question(s). 
What results from these specific statistical methods are needed to support your hypothesized answer?)
