---
title: "Final Report"
author: "Isabel Siebrecht, Franklin Feng, & Kaylin Hernandez"
date: "11/16/21"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: 'due November 16, 2021 by 11:59 PM '
---

```{r load-packages, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(infer)
library(tidymodels)
install.packages("ggpubr")
library("ggpubr")

knitr::opts_chunk$set(eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
``` 

```{r load-data, echo=FALSE}
hdi1 <- read.csv("../data/HDI_Table1.csv")
hdi4 <- read.csv("../data/HDI_Table4.csv")
hdi5 <- read.csv("../data/HDI_Table5.csv")
hdi8 <- read.csv("../data/HDI_Table8.csv")
hdi10 <- read.csv("../data/HDI_Table10.csv")
adolescent_births <- read.csv("../data/Adolescent_Births.csv")
contraceptive_prevalence <- read.csv("../data/Contraceptive_Prevalence.csv")
```

```{r clean-data, echo=FALSE}
hdi1_clean <- slice(hdi1, -c(1:7, 74, 128, 166, 200, 207:270))
hdi1_clean <- select(hdi1_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11))
names(hdi1_clean) <- c("HDI Rank", "Country", "HDI 2019", "Life Expectancy Birth", "Expected Years Schooling", "Mean Years Schooling", "GNI Per Capita", "GNI Per Capita Minus HDI", "HDI 2018")

hdi4_clean <- slice(hdi4, -c(1:7, 74, 128, 166, 200, 207:264))
hdi4_clean <- select(hdi4_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23))
names(hdi4_clean) <- c("HDI Rank", "Country", "Gender Development Index", "Gender Development Index Group", "HDI Female", "HDI Male", "Life Expectancy Birth Female", "Life Expectancy Birth Male", "Expected Years Schooling Female", "Expected Years Schooling Male", "Mean Years Schooling Female", "Mean Years Schooling Male", "Estimated GNI per Capita Female", "Estimated GNI per Capita Male")

hdi5_clean <- slice(hdi5, -c(1:7, 74, 128, 166, 200, 207:270))
hdi5_clean <- select(hdi5_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17))
names(hdi5_clean) <- c("HDI Rank", "Country", "Gender Inequality Index", "Gender Inequality Index Rank", "Maternal Mortality Ratio", "Adolescent Birth Rate", "Seats Parliament Female", "Secondary Education Female", "Secondary Education Male", "Labour Participation Female", "Labour Participation Male")

hdi8_clean <- slice(hdi8, -c(1:7, 74, 128, 166, 200, 207:264))
hdi8_clean <- select(hdi8_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23, X.25, X.27, X.29 : X.33))
names(hdi8_clean) <- c("HDI Rank", "Country", "Infants Breastfed", "Lack DTP Immunization", "Lack Measles Immunization", "Child Malnutrition", "Mortality Rate Infant", "Mortality Rate Under 5", "Mortality Rate Female", "Mortality Rate Male", "Mortality Rate Noncommunicable Female", "Mortality Rate Noncommunicable Male", "Malaria Incidence", "TB Incidence", "HIV Prevalence", "Healthy Life Expectancy at Birth", "Health Expenditure of GDP")

hdi10_clean <- slice(hdi10, -c(1:7, 74, 128, 166, 200, 207:260))
hdi10_clean <- select(hdi10_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23))
names(hdi10_clean) <- c("HDI Rank", "Country", "GDP", "GDP per Capita", "GDP Annual Growth", "Labour Share", "Gross Fixed Capital Formation of GDP", "General Government Consumption of GDP", "General Government Consumption Annual Growth", "Total Tax Revenue of GDP", "Taxes Income Profits Capital Gains of TTR", "External Debt Stock of GNI", "Total Debt Service of GNI", "Consumer Price Index")

adolescent_births_clean <- select(adolescent_births, select = c(ParentLocation, Location, Period, IsLatestYear, FactValueNumeric, Value))
names(adolescent_births_clean) <- c("World Region", "Country", "Period", "Is Latest Year", "Adolescent Births", "Adolescent Births Text")

contraceptive_prevalence_clean <- slice(contraceptive_prevalence, -c(1:8))
contraceptive_prevalence_clean <- select(contraceptive_prevalence_clean, select = c(X, X.2, X.3, X.4, X.5, X.6, X.19))
names(contraceptive_prevalence_clean) <- c("Country", "Survey Start Year", "Survey End Year", "Age Group", "Contraceptive Any Method", "Contraceptive Modern Method", "Contraceptive Traditional Method")
```

```{r join-datasets, echo=FALSE}

country_joined <- hdi1_clean %>%
  inner_join(hdi10_clean, by = "Country")%>%
  inner_join(hdi4_clean, by = "Country")%>%
  inner_join(hdi5_clean, by = "Country")%>%
  inner_join(hdi8_clean, by = "Country")

contraceptives_joined <- contraceptive_prevalence_clean %>%
  inner_join(adolescent_births_clean, by = "Country")
```

## Introduction

Our group chose to research what influences the rate of teenage pregnancies because we consider premature pregnancies one of the most significant indicators of a country’s development and prosperity. Considering the repercussions of lost economic contributions and increased dependency on costly public programs [https://www.ncbi.nlm.nih.gov/books/NBK219229/], we hope our findings can encourage governments to reduce the rate of teenage pregnancies by focusing on the strongest predictor variables we identify in this report.  

In essence, we attempt to answer how a country’s quality of life, measured by human and economic resources, correlates to the birth rates for teenagers aged 15 to 19 years old. To answer our question, our explanatory variables of adolescent pregnancy rates are female educational attainment, GDP annual growth rate, life expectancy at birth, HDI, labor force participation rate, and gender inequality index.  
Information provided by Youth.gov supports our predictors because they broadly encompass the individual, family, and community that influence teenage pregnancies ([Youth.gov](https://youth.gov/youth-topics/pregnancy-prevention/risk-and-protective-factors)).  

We pulled our data on adolescent birth rate from the World Health Organization who collected the results via various national and international surveys and country reports. We also pulled data from the 2020 Human Development Report provided by the United Nations. This data consisted of female educational attainment (number of years of schooling), GDP, life expectancy at birth, HDI, labor force participation rate, and gender inequality index. The gender equality index measures the gender disparity in a country on a scale from 0 to 1. The greater the value, the greater the measure of inequality.  

As a disclaimer, we acknowledge that our findings are based on preconceived, Western assumptions of what qualifies as a global indicator of “status.” Thus, we conducted this report on our assumptions of what defines a country’s prosperity and development.  

## Methodology
## Intro to data
To allow the reader a complete understanding of our findings, we first focused on analyzing our main variable of interest, adolescent birth rates. Utilizing data from 195 countries, we produced a density plot and corresponding quantile values below. We would like to note that throughout the report adolescent birth rate is measured as the number of births per 1000 women aged 15-19 years old in a given population.

```{r total-rate, fig.height=3, fig.width=7, echo=FALSE, warning=FALSE}
country_joined %>%
  ggplot(aes(x = as.numeric(`Adolescent Birth Rate`))) +
  geom_boxplot() +
  labs(title= "Adolescent Birth Rates Worldwide", subtitle = "Averaged Over a 5 Year Period From 2015-2020",
       x= "Adolescent Birth Rate (per 1,000 Women Aged 15-19)")
```

```{r quantile, echo = FALSE, warning=FALSE, message=FALSE, error=FALSE}
quantile_set <- country_joined %>%
  select(`Adolescent Birth Rate`) %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`))

  quantile(quantile_set$Adolescent_Birth_Rate, c(0, 0.25, 0.50, 0.75, 1))
```

We see that the middle 50% of countries have a teenage birth rate between 13.2 and 70.5. Considering a low birth rate is 10-20 births per 1,000 women and a high birth rate is 40-50 births per 1,000 women ([Child Trends](https://web.archive.org/web/20160526190803/http://www.childtrends.org/?indicators=fertility-and-birth-rates)), our box plot shows a relatively broad distribution rates. Thus, this right-skewed distribution can allow us to compare indicators on both sides of the spectrum. We also have a few outlines, which are Chad (161.1), Mali (169.1), and Niger (186.5). All three have HDI ranks 187, 184, and 189, respectively.

>>> If possible I will fix to eliminate y-axis values + add subtitle
>> was it actually averaged?? I thought our data was only from 2019??


## Relationships

After looking at the distribution of birth rate data, we decided to explore if there was a relationship between our chosen factors and adolescent pregnancy rates. Therefore, we plotted our six explanatory variables (maternal mortality rate, female educational attainment, GDP, life expectancy at birth, 2019 HDI index, and gender development index) onto separate scatter plots. The relationships will then be modeled with a linear regression to determine the significance of adolescent birth rates an independent indicator has. 

**Is Maternal Mortality Ratio Related to Adolescent Birth Rate?**

```{r maternal-mortality, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE, fig.height=3, fig.width=7}
ggplot(country_joined, aes(x = as.numeric(`Maternal Mortality Ratio`), 
                           y = as.numeric(`Adolescent Birth Rate`))) +
  geom_point(alpha = 0.6) +
  labs(title= "Does Maternal Mortality Ratio Affect Adolescent Birth Rate?", subtitle = "Observations From 185 Countries",
       x = "Maternal Mortality Ratio", y = "Adolescent Birth Rate")

```
```{r cor-maternal-mortality, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Maternal Mortality Ratio`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use = "complete.obs")

```

The data shows a positive, linear trend, and we see that as maternal mortality increases, so does adolescent birth rate across countries. The maternal mortality ratio signifies the number of maternal deaths per 100,000 live births during the same time. We expect countries with high adolescent birth rates to have high maternal mortality ratios. The relationship between the two variables is positive and relatively strong, considering the correlation coefficient is 0.7314.  

**Is Female Educational Attainment Related to Adolescent Birth Rate?**

```{r fem-education, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  filter(!`Mean Years Schooling Female` == "..") %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`)) %>%
  mutate(Mean_Years_Schooling_Female = as.numeric(`Mean Years Schooling Female`)) %>%
  
  ggplot(aes(x= Mean_Years_Schooling_Female, y= Adolescent_Birth_Rate)) +
  geom_point(alpha=0.6) +
  labs(
    title= "Does Female Educational Attainment Affect Adolescent Birth Rate?",
    x="Mean Years Schooling for Women",
    y="Adolescent Birth Rate"
  )
```
```{r cor-female-schooling, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Mean Years Schooling Female`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```

This graph highlights a relatively strong negative correlation (r = -0.7249) that for a given country, as the mean years of schooling for women increases, the rate of adolescent births decreases. This logically makes sense because the increase in time spent in school also raises the likelihood of women finding a career in the workforce along with being actively informed about sexual health and safe practices. These factors directly contribute to an overall lower adolescent birth rate.  

**Is GDP Annual growth rate related to adolescent birth rate?**

```{r GDP, echo=FALSE, error=FALSE, fig.height=3, fig.width=7, message=FALSE, warning=FALSE}
ggplot(country_joined, aes(x = as.numeric(`GDP Annual Growth`), 
                           y = as.numeric(`Adolescent Birth Rate`), 
                           color = as.numeric(`HDI Rank.x`))) +
  geom_point() +
  geom_smooth() +
  labs(
    title= "Is GDP Annual Growth Rate Related to Adolescent Birth Rates?",
    x="GDP Annual Growth Rate",
    y="Adolescent Birth Rate",
    color = "HDI Rank") 
```
```{r cor-GDP, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`GDP Annual Growth`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
Gross Domestic Product (GDP) annual growth rate signifies the yearly average rate of change of GDP at market prices based on constant local currency, for a given national economy, during a specified time ([The Balance](https://www.thebalance.com/u-s-gdp-growth-3306008)). This graph shows an unclear relationship between GDP annual growth rate and adolescent birth rate (per 1,000 15-19-year-olds)--on neither the regular or log scale. This was surprising to see because we assumed countries with lower GDP annual growths would have higher adolescent birth rates. To double-check, we calculated the correlation coefficient and found a very weak, negative relationship (r=-0.1261). However, because there is still a slight correlation and we need an economic variable to compare, we will still include GDP annual growth in our linear regression.  

**Is life expectancy at birth related to adolescent birth rate?**
```{r life-exp, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
ggplot(country_joined, aes(x = as.numeric(`Life Expectancy Birth`), 
                           y = as.numeric(`Adolescent Birth Rate`))) +
  geom_point(alpha=0.6)+
  labs(
    title= "Is Life Expectacy at Birth Related to Adolescent Birth Rate?",
    x="Life Expectacy at Birth",
    y="Adolescent Birth Rate"
  )

```
```{r cor-life-exp, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Life Expectancy Birth`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
This graph shows a strong, negative, linear relationship (r=-0.8003) between life expectancy at birth and adolescent birth rate (per 1,000 women aged 15-19 years old), meaning as life expectancy increases, teenage birth rate decreases. This makes sense because women would have less need to reproduce if their children are expected to live for a long time.      

**Is HDI in 2019 related to adolescent birth rate?**
```{r HDI, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  ggplot(aes(x= as.numeric(`HDI 2019`), 
             y= as.numeric(`Adolescent Birth Rate`))) +
  geom_point(alpha=0.6) +
  labs(
    title= "Is HDI in 2019 Related to Adolescent Birth Rate?",
    x="2019 HDI",
    y="Adolescent Birth Rate"
  )
```
```{r cor-HDI-2019, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`HDI 2019`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
HDI, also known as the Human Development Index, is a statistic used to rank countries. This plot shows a strong, negative, linear relationship (r=-0.8196) between 2019 HDI and adolescent birth rate (per 1,000 women aged 15-19 years old), meaning as the HDI increases, the rate of adolescent birth rates decrease. This relationship makes sense because HDI is an indicator of wealth, and the more money generally correlates with more contraceptives, abortions, better maternal and child healthcare, and .  

**Is labor force participation rate related to adolescent birth rate?**
```{r labor-force, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  ggplot(aes(x= as.numeric(`Labour Participation Female`), 
             y= as.numeric(`Adolescent Birth Rate`)))+
  geom_point(alpha=0.5) +
  labs(
    title= "Is Female Labor Participation Related to Adolescent Birth Rate?",
    x="Female Labor Participation",
    y="Adolescent Birth Rate"
  )
```
```{r cor-labor-participation, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Labour Participation Female`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
From the plot, we only somewhat see a relationship between female labor participation and adolescent birth rate, and our correlation coefficient of 0.3103 confirms a weak, positive linear correlation. Roughly, as female labor participation increases, the birth rate per 1,000 women aged 15-19 years old also increases. This is slightly surprising because we predicted that countries with high teenage birth rates would have higher female labor participation levels. We assumed the young women would be at home, mainly because we saw this earlier with female educational attainment.

**Is the gender develop index related to adolescent birth rate?**
```{r gender-inequality, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  ggplot(aes(x= as.numeric(`Gender Development Index`), 
             y= as.numeric(`Adolescent Birth Rate`))) +
  geom_point(alpha=0.6) +
  labs(
    title= "Is Gender Development Index Related to Adolescent Birth Rate?",
    x="Gender Development Index",
    y="Adolescent Birth Rate"
  )
```
```{r cor-GDI, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Gender Development Index`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
The gender development index measures gender inequality through gender-based gaps in education, life expectancy, and incomes, and the closer the value is to 1, the more equal a society is. We see that as the gender development index approaches 1, there is a slight linear decrease (r= -0.5148) in adolescent birth rate (per 1,000 15-19-year-olds). This relationship makes sense because we assume that more equal societies would allow women to make their own reproductive choices. They are also more likely to have resources and healthcare options for women.   

# Linear regression
After looking to see possible relationships, we decided to run a linear regression analysis in order to determine which predictors had the most influence over teenage pregnancy rates and what that relationship is. The general null hypothesis test for every explanatory variable is beta# = 0, which indicates that there is no difference in influence when we control for each one individually. 

**Which variable has the most significant correlation to adolescent birth rate?**
```{r total-linear-regression}
country_joined <- country_joined %>%
  filter(!`GDP` == "..") %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  filter(!`Maternal Mortality Ratio` == "..") %>%
  filter(!`Mean Years Schooling Female` == "..") %>%
  filter(!`Labour Participation Female` == "..") %>%
  filter(!`Gender Inequality Index` == "..") %>%
  mutate(GDP = as.numeric(`GDP`)) %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`)) %>%
  mutate(Country_new = as.factor(`Country`)) %>%
  mutate(`Maternal Mortality Ratio` = str_replace(`Maternal Mortality Ratio`, ",", "")) %>%
  mutate(Maternal_Mortality_Ratio = as.numeric(`Maternal Mortality Ratio`)) %>%
  mutate(Mean_Years_Schooling_Female = as.numeric(`Mean Years Schooling Female`)) %>%
  mutate(Life_Expectancy_Birth = as.numeric(`Life Expectancy Birth`)) %>%
  mutate(HDI_2019 = as.numeric(`HDI 2019`)) %>%
  mutate(Labour_Participation_Female = as.numeric(`Labour Participation Female`)) %>%
  mutate(Gender_Inequality_Index = as.numeric(`Gender Inequality Index`))

linreg_total <- country_joined %>%
  select(Country_new, Adolescent_Birth_Rate, GDP, Maternal_Mortality_Ratio, Mean_Years_Schooling_Female, Life_Expectancy_Birth, HDI_2019, Labour_Participation_Female, Gender_Inequality_Index)
         
mainfit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Adolescent_Birth_Rate ~ GDP + Maternal_Mortality_Ratio + Mean_Years_Schooling_Female + Life_Expectancy_Birth + HDI_2019 + Labour_Participation_Female + Gender_Inequality_Index , data = linreg_total)
tidy(mainfit)

#Country_new is available to be added, but produces N/A values

#The 2 most statistically significant variables are Labour Participation Female and Gender Inequality Index. Both fall under the p = 0.05 range
```

## Interpretation
> which variable(s) was rejected by the null hypothesis by having no difference in influence over birth rate
> which variable(s) led us to fail to reject the null hypothesis and thus DOES have a difference in influence over birth rate. 
>> what is this relationship/interpret
> Where do we go from here?

```{r GDP-linreg, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
country_joined <- country_joined %>%
  filter(!`GDP` == "..") %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  #filter(!`GNI Per Capita` == "..") %>%
  filter(!`GDP per Capita` == "..") %>%
  mutate(GDP = as.numeric(`GDP`)) %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`)) %>%
  mutate(HDI_Rank = as.factor(`HDI Rank.x`)) %>%
  #mutate(GNI_Per_Capita = as.numeric(`GNI per Capita`)) %>%
  mutate(`GDP per Capita` = str_replace(`GDP per Capita`, ",", "")) %>%
  mutate(GDP_Per_Capita = as.numeric(`GDP per Capita`))
  

gdp_linreg <- country_joined %>%
  select(Adolescent_Birth_Rate, GDP, HDI_Rank, GDP_Per_Capita)
         
mainfit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Adolescent_Birth_Rate ~ GDP + HDI_Rank + GDP_Per_Capita, data = gdp_linreg)

```

```{r HDI Grouping, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
cat_new <- country_joined %>%
  mutate(HDI_final = as.numeric(`HDI Rank.x`)) %>%
  mutate(HDI_cat = cut(
    HDI_final,
    breaks = c(-Inf, 66, 119, 156, 189, Inf),
    labels = c(
      "Very High", "High", "Medium", "Low", "Other")))
```

# Looking further at Gender Inequality Index
```{r}
ggplot(data = cat_new) +
  geom_point(aes(x = as.numeric(`Gender Inequality Index`), y = as.numeric(`Adolescent_Birth_Rate`), color = HDI_cat)) +
  labs(x = "Gender Inequality Index", 
       y = "Adolescent Birth Rate",
       color = "HDI Category")
```
HDI reflects economic prosperity in a country, so we broke the HDI rankings up into four categories to visualize the hypothesis that wealth plays a major roll in access to equality. Since gender inequality is such a broad issue we wanted to look at wealth desperation in conjunction. This plot supports the idea that wealthier countries tend to have lower gender equality disparities and lower rates of teen births, and vice versa. Our plot, organized by HDI rank, can help us select which countries will benefit most from our intervention targeting wealth and gender inequality to improve adolescent birth rates.

# Summary & Next Steps
> referencing the importance of teenage birth rate and our research question, 
>> just summarize our the relationships we found and which variable we found to be the most significant
> also explain why this correlation is important and how researchers/governments can use this knowledge to reduce birth rates
> know correlation doesn't mean causation
> if you reference international organizations, note that program implementations cannot be generalized acorss the globe based on this data, which is simply a global sample from 2019. Each country (and maybe disaggregate further) needs further reseach and separate analysis.

