---
title: "Final Report"
author: "Isabel Siebrecht, Franklin Feng, & Kaylin Hernandez"
date: "11/16/21"
output:
  pdf_document: default
  html_document:
    df_print: paged
subtitle: 'due November 16, 2021 by 11:59 PM '
---

```{r load-packages, echo = FALSE, error = FALSE, message = FALSE, warning = FALSE}
library(tidyverse)
library(dplyr)
library(ggplot2)
library(tidyr)
library(readr)
library(infer)
library(tidymodels)
install.packages("ggpubr")
library("ggpubr")

knitr::opts_chunk$set(eval = TRUE, echo = FALSE, message = FALSE, warning = FALSE)
``` 

```{r load-data, echo=FALSE}
hdi1 <- read.csv("../data/HDI_Table1.csv")
hdi4 <- read.csv("../data/HDI_Table4.csv")
hdi5 <- read.csv("../data/HDI_Table5.csv")
hdi8 <- read.csv("../data/HDI_Table8.csv")
hdi10 <- read.csv("../data/HDI_Table10.csv")
adolescent_births <- read.csv("../data/Adolescent_Births.csv")
contraceptive_prevalence <- read.csv("../data/Contraceptive_Prevalence.csv")
```

```{r clean-data, echo=FALSE, message=FALSE, warning=FALSE}
hdi1_clean <- slice(hdi1, -c(1:7, 74, 128, 166, 200, 207:270))
hdi1_clean <- select(hdi1_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11))
names(hdi1_clean) <- c("HDI Rank", "Country", "HDI 2019", "Life Expectancy Birth", "Expected Years Schooling", "Mean Years Schooling", "GNI Per Capita", "GNI Per Capita Minus HDI", "HDI 2018")

hdi4_clean <- slice(hdi4, -c(1:7, 74, 128, 166, 200, 207:264))
hdi4_clean <- select(hdi4_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23))
names(hdi4_clean) <- c("HDI Rank", "Country", "Gender Development Index", "Gender Development Index Group", "HDI Female", "HDI Male", "Life Expectancy Birth Female", "Life Expectancy Birth Male", "Expected Years Schooling Female", "Expected Years Schooling Male", "Mean Years Schooling Female", "Mean Years Schooling Male", "Estimated GNI per Capita Female", "Estimated GNI per Capita Male")

hdi5_clean <- slice(hdi5, -c(1:7, 74, 128, 166, 200, 207:270))
hdi5_clean <- select(hdi5_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17))
names(hdi5_clean) <- c("HDI Rank", "Country", "Gender Inequality Index", "Gender Inequality Index Rank", "Maternal Mortality Ratio", "Adolescent Birth Rate", "Seats Parliament Female", "Secondary Education Female", "Secondary Education Male", "Labour Participation Female", "Labour Participation Male")

hdi8_clean <- slice(hdi8, -c(1:7, 74, 128, 166, 200, 207:264))
hdi8_clean <- select(hdi8_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23, X.25, X.27, X.29 : X.33))
names(hdi8_clean) <- c("HDI Rank", "Country", "Infants Breastfed", "Lack DTP Immunization", "Lack Measles Immunization", "Child Malnutrition", "Mortality Rate Infant", "Mortality Rate Under 5", "Mortality Rate Female", "Mortality Rate Male", "Mortality Rate Noncommunicable Female", "Mortality Rate Noncommunicable Male", "Malaria Incidence", "TB Incidence", "HIV Prevalence", "Healthy Life Expectancy at Birth", "Health Expenditure of GDP")

hdi10_clean <- slice(hdi10, -c(1:7, 74, 128, 166, 200, 207:260))
hdi10_clean <- select(hdi10_clean, select = -c(X.1, X.3, X.5, X.7, X.9, X.11, X.13, X.15, X.17, X.19, X.21, X.23))
names(hdi10_clean) <- c("HDI Rank", "Country", "GDP", "GDP per Capita", "GDP Annual Growth", "Labour Share", "Gross Fixed Capital Formation of GDP", "General Government Consumption of GDP", "General Government Consumption Annual Growth", "Total Tax Revenue of GDP", "Taxes Income Profits Capital Gains of TTR", "External Debt Stock of GNI", "Total Debt Service of GNI", "Consumer Price Index")

adolescent_births_clean <- select(adolescent_births, select = c(ParentLocation, Location, Period, IsLatestYear, FactValueNumeric, Value))
names(adolescent_births_clean) <- c("World Region", "Country", "Period", "Is Latest Year", "Adolescent Births", "Adolescent Births Text")

contraceptive_prevalence_clean <- slice(contraceptive_prevalence, -c(1:8))
contraceptive_prevalence_clean <- select(contraceptive_prevalence_clean, select = c(X, X.2, X.3, X.4, X.5, X.6, X.19))
names(contraceptive_prevalence_clean) <- c("Country", "Survey Start Year", "Survey End Year", "Age Group", "Contraceptive Any Method", "Contraceptive Modern Method", "Contraceptive Traditional Method")
```

```{r join-datasets, echo=FALSE, message=FALSE, warning=FALSE}

country_joined <- hdi1_clean %>%
  inner_join(hdi10_clean, by = "Country")%>%
  inner_join(hdi4_clean, by = "Country")%>%
  inner_join(hdi5_clean, by = "Country")%>%
  inner_join(hdi8_clean, by = "Country")

contraceptives_joined <- contraceptive_prevalence_clean %>%
  inner_join(adolescent_births_clean, by = "Country")
```

## Introduction

In this project, we chose to research what factors in our world influences the 
rate of teenage pregnancies. We consider premature pregnancies to be a very 
significant indicator of a country’s development and prosperity. Considering the
repercussions of lost economic contributions and dependency on costly public 
programs [https://www.ncbi.nlm.nih.gov/books/NBK219229/], we hope our findings 
can spur a decrease in the rate of teenage pregnancies by encouraging 
governments to focus on the strongest predictor variables linked to adolescent 
births.  

In essence, we attempt to answer how a country’s quality of life, measured by 
human and economic resources, correlates to the birth rates for teenagers aged 
15 to 19 years old. To answer our question, our explanatory variables of 
adolescent pregnancy rates are female educational attainment, GDP annual growth 
rate, life expectancy at birth, HDI, labor force participation rate, and gender 
inequality index. 

Information provided by Youth.gov supports our predictors because they broadly 
encompass the individual, family, and community that influence teenage 
pregnancies ([Youth.gov](https://youth.gov/youth-topics/pregnancy-prevention/risk-and-protective-factors)).  

As a note, we acknowledge that our findings are based on preconceived, Western 
assumptions of what qualifies as a global indicator of “status.” Thus, we 
conducted this report on our assumptions of what defines a country’s prosperity 
and development.

  
## Methodology

There were a few sources that our data was obtained. We mainly worked with the 
2020 Human Development Report, courtesy of the United Nationas Development 
Programme. With a vast array of variables available to analyze, we selected 
female educational attainment (number of years of schooling), GDP and related 
economic indicators, adolescent birth rate, life expectancy at birth, maternal 
mortality rate, HDI index, labor force participation rate, and the gender 
inequality index. We also obtained additional data on adolescent birth rates 
from the World Health Organization who collected the results via various 
national and international surveys and country reports, though this dataframe 
was largely utilized for reference purposes.

The logic of constructing our project was to try and explore different 
relationships between these variables and try to create combinations that lead 
back to their effect(s) on adolescent birth rates. There are many different ways
of exploring this data and to create the clearest, most straightforward analysis
we created various plots and analyses. Ultimately, we seek to maximize our 
explanatory power in the relationship between a country's resources and 
adolescent birth rates and identifying those specific variables that really move
the needle.

## Intro to data
  
To allow the reader a complete understanding of our findings, we first focused 
on analyzing our main variable of interest, adolescent birth rates. Utilizing 
data from 195 countries, we produced a density plot and corresponding quantile 
values below. We would like to note that throughout the report adolescent birth 
rate is measured as the number of births per 1000 women aged 15-19 years old in 
a given population.  

```{r total-rate, fig.height=3, fig.width=7, echo=FALSE, warning=FALSE}
country_joined %>%
  ggplot(aes(x = as.numeric(`Adolescent Birth Rate`))) +
  geom_boxplot() +
  labs(title= "Adolescent Birth Rates Worldwide", subtitle = "Collected in 2019",
       x= "Adolescent Birth Rate (per 1,000 Women Aged 15-19)")
```

```{r quantile, echo = FALSE, warning=FALSE, message=FALSE, error=FALSE}
quantile_set <- country_joined %>%
  select(`Adolescent Birth Rate`) %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`))

  quantile(quantile_set$Adolescent_Birth_Rate, c(0, 0.25, 0.50, 0.75, 1))
```

We see that the middle 50% of countries have a teenage birth rate between 13.4 
and 72.9. Considering a low birth rate is 10-20 births per 1,000 women and a 
high birth rate is 40-50 births per 1,000 women ([Child Trends](https://web.archive.org/web/20160526190803/http://www.childtrends.org/?indicators=fertility-and-birth-rates)),
our box plot shows a relatively broad distribution rate. Thus, this right-skewed
distribution can allow us to compare indicators on both sides of the spectrum. 
We also have a few outlines, which are Chad (161.1), Mali (169.1), and Niger 
(186.5). All three have HDI ranks 187, 184, and 189, respectively.  
  

## Relationships

After looking at the distribution of birth rate data, we decided to explore if there was a relationship between our chosen factors and adolescent pregnancy rates. Therefore, we plotted our seven explanatory variables (maternal mortality rate, female educational attainment, GDP, life expectancy at birth, 2019 HDI index, and gender development index) onto separate scatter plots. The relationships will then be modeled with a linear regression to determine the significance of adolescent birth rates an independent indicator has.  
  
  
**Is Maternal Mortality Ratio Related to Adolescent Birth Rate?**
  
```{r maternal-mortality, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE, fig.height=3, fig.width=7}
mortalities <- country_joined %>%
  mutate(Maternal_Mortality_Adj = as.numeric(`Maternal Mortality Ratio`) / 100) %>%
  filter(!is.na(Maternal_Mortality_Adj))

ggplot(mortalities, aes(x = Maternal_Mortality_Adj, 
                           y = as.numeric(`Adolescent Birth Rate`), color = as.numeric(`Mortality Rate Infant`))) +
  geom_point(alpha = 0.6) +
  labs(title= "Does Maternal Mortality Ratio Affect Adolescent Birth Rate?", 
       subtitle = "Per 1,000 Live Births",
       x = "Maternal Mortality Ratio", y = "Adolescent Birth Rate", color = "Infant Mortality Rate")
```

```{r cor-maternal-mortality, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Maternal Mortality Ratio`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", 
    use = "complete.obs")

```
  
The data shows a positive, linear trend, and we see that as maternal mortality 
increases, so does adolescent birth rate across countries. We expect countries 
with higher adolescent birth rates to have high maternal mortality ratios, which
can be due to lack of access to modern healthcare among other factors. The 
relationship between the two variables is positive and relatively strong, 
considering the correlation coefficient is 0.731. Additionally, we overlaid 
another variable of infant mortality rates for each of the countries. We clearly
see that this trend continues to have this relationship of higher maternal 
mortality rates and adolescent birth rates also contributes to higher infant 
mortality rates. 

**Is Female Educational Attainment Related to Adolescent Birth Rate?**

```{r fem-education dataframe, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
fem_education <- country_joined %>%
  select(`Adolescent Birth Rate`, `Mean Years Schooling Female`, `Mean Years Schooling Male`, `Secondary Education Female`, `Secondary Education Male`) %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  filter(!`Mean Years Schooling Female` == "..") %>%
  filter(!`Secondary Education Female` == "..") %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`)) %>%
  mutate(Mean_Years_Schooling_Female = as.numeric(`Mean Years Schooling Female`)) %>%
  mutate(Mean_Years_Schooling_Male = as.numeric(`Mean Years Schooling Male`)) %>%
  mutate(Secondary_Education_Female = as.numeric(`Secondary Education Female`)) %>%
  mutate(Secondary_Education_Male = as.numeric(`Secondary Education Male`))
```

```{r fem-education plot, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
ggplot(data = fem_education, mapping = aes(x= Mean_Years_Schooling_Female, y= Adolescent_Birth_Rate)) +
  geom_col(fill = "Skyblue") +
  geom_smooth() +
  labs(
    title= "Does Female Educational Attainment Affect Adolescent Birth Rate?",
    subtitle = "Per 1,000 Live Births",
    x="Mean Years Schooling for Women",
    y="Adolescent Birth Rate"
  )
```


```{r cor-female-schooling, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Mean Years Schooling Female`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")
```
  
This graph highlights a relatively strong negative correlation (r = -0.7249) 
that for a given country, as the mean years of schooling for women increases, 
the rate of adolescent births decreases. This logically makes sense because the 
increase in time spent in school also raises the likelihood of women finding a 
career in the workforce along with being actively informed about sexual health 
and safe practices. These factors directly contribute to an overall lower 
adolescent birth rate.


**Is GDP Annual growth rate related to adolescent birth rate?**

```{r GDP-dataset}
gdp_data <- country_joined %>%
  select(`GDP`, `GDP Annual Growth`, `GDP per Capita`, `Consumer Price Index`, `GNI Per Capita`, `HDI Rank.x`, `Adolescent Birth Rate`) %>%
  filter(!`GDP` == "..") %>%
  filter(!`Consumer Price Index` == "..") %>%
  filter(!`GNI Per Capita` == "..") %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  mutate(`GDP per Capita` = str_replace(`GDP per Capita`, ",", "")) %>%
  mutate(`GDP` = str_replace(`GDP`, ",", "")) %>%
  mutate(`GNI Per Capita` = str_replace(`GNI Per Capita`, ",", "")) %>%
  mutate(`Consumer Price Index` = str_replace(`Consumer Price Index`, ",", "")) %>%
  mutate(GDP_num = as.numeric(`GDP`)) %>%
  mutate(GDP_Annual_Growth = as.numeric(`GDP Annual Growth`)) %>%
  mutate(GDP_per_Capita = as.numeric(`GDP per Capita`)) %>%
  mutate(Consumer_Price_Index = as.numeric(`Consumer Price Index`)) %>%
  mutate(GNI_Per_Capita = as.numeric(`GNI Per Capita`)) %>%
  #mutate(HDI_Rankx = as.numeric(`HDI Rank.x`)) %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`))
```

  
```{r GDP, echo=FALSE, error=FALSE, fig.height=3, fig.width=7, message=FALSE, warning=FALSE}
ggplot(gdp_data, aes(x = as.numeric(`GDP Annual Growth`), 
                           y = as.numeric(`Adolescent Birth Rate`), 
                           color = as.numeric(`HDI Rank.x`))) +
  geom_point() +
  labs(
    title= "Does GDP Annual Growth Rate Affect Adolescent Birth Rates?",
    subtitle = "per 1000 Live Births",
    x="GDP Annual Growth Rate",
    y="Adolescent Birth Rate",
    color = "HDI Rank")
```

```{r cor-GDP, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`GDP Annual Growth`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
  
```{r GDP-linreg}
mainfit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Adolescent_Birth_Rate ~ GDP_num + GDP_Annual_Growth + GDP_per_Capita + Consumer_Price_Index +  GNI_Per_Capita, data = gdp_data)
tidy(mainfit, conf.int=FALSE)
```
  
We next asked ourselves if strictly economic indicators can be telling in 
differences in adolescent birth rate. The results were much more muddy than 
anticipated. For example, we plotted Gross Domestic Product (GDP) annual growth 
rate, which signifies the yearly average rate of change of GDP at market prices 
based on constant local currency, for a given national economy, during a 
specified time ([The Balance](https://www.thebalance.com/u-s-gdp-growth-3306008)).
The plot shows an unclear relationship between GDP annual growth rate and 
adolescent birth rate. It is to be noted, however, that the HDI index has a 
substantial impact on the rate of adolescent births as clearly shown on the plot
delineated by color.

We further explored the relationships with economic variables such as GDP, 
Consumer Price Index, GNI (Gross National Income) per Capita using a linear 
regression model to seek if economic indicators on their own can tell about 
adolescent birth rates in a given country. GDP Annual Growth and GNI per Capita 
were the two economic indicators that were the most significant in the model, 
though both failed to reach the appropriate p-level threshold to be counted as 
statistically significant.

Finally, we calculated the correlation coefficient and found a very weak, 
negative relationship (r = -0.1261). We can therefore conclude that just by 
looking at a given country's economic condition and growth is a fairly poor 
indicator of the adolescent birth rate. This definitely makes sense since there
are many developing nations with high GDP growth but still lack advanced 
infrastructure for healthcare, while there are low growth GDP countries that 
have reached the post modern classification and have highly robust healthcare
access.  

  
**Is life expectancy at birth related to adolescent birth rate?**

```{r life-exp, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
ggplot(country_joined, aes(x = as.numeric(`Life Expectancy Birth`), 
                           y = as.numeric(`Adolescent Birth Rate`), color = as.numeric(`Health Expenditure of GDP`))) +
  geom_point(alpha=0.8)+
  labs(
    title= "Does Life Expectancy at Birth Affect Adolescent Birth Rate?",
    subtitle = "Mean Expected Years for a Given Country",
    x="Life Expectancy at Birth",
    y="Adolescent Birth Rate", color = "Health Expenditure of GDP"
  )

```

```{r cor-life-exp, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Life Expectancy Birth`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
  
This graph shows a strong, negative, linear relationship (r = -0.778) between 
life expectancy at birth and adolescent birth rate, meaning as life expectancy 
increases, adolescent birth rate decreases. This makes logical sense because 
women would have less need to reproduce if their children are expected to have a
long lifespan. In addition, we applied an overlay of Health Expenditure of GDP 
to study if this variable affects both life expectancy and adolescent birth 
rates. As seen, there is not an extremely clear pattern exhibited, but in 
general the trend tends to be that those countries with a higher devotion of 
Health Expenditure relative to their GDP came out to longer life expectancies 
and lower adolescent birth rates.
  
  
**Is HDI in 2019 related to adolescent birth rate?**

```{r cor-HDI-2019, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`HDI 2019`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
  
HDI, also known as the Human Development Index, is a statistic used to rank 
countries. There is clearly a relatively strong negative linear relationship 
(r = -0.774) between 2019 HDI and adolescent birth rate (per 1,000 women aged 
15-19 years old), meaning as the HDI increases, the rate of adolescent birth 
rates decrease. This relationship makes sense because unlike that of other 
economic indicators like GDP and GNI above, HDI is a more holistic indicator of 
a country's resources, its condition both economically and health-wise, and the 
level of democracy that exists. The confluence of these factors helps to build 
the HDI index and therefore it commands a far better relationship with 
adolescent birth rates than many other single standing variables.


**Is labor force participation rate related to adolescent birth rate?**
  
```{r labor-force, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  ggplot(aes(x= as.numeric(`Labour Participation Female`), 
             y= as.numeric(`Adolescent Birth Rate`)))+
  geom_point(alpha=0.5) +
  labs(
    title= "Does Female Labor Participation Affect Adolescent Birth Rate?",
    x="Female Labor Participation",
    y="Adolescent Birth Rate"
  )
```

```{r cor-labor-participation, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Labour Participation Female`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```
  
From the plot, we only somewhat see a relationship between female labor 
participation and adolescent birth rate, and our correlation coefficient of 
0.3103 confirms a weak, positive linear correlation. Roughly, as female labor 
participation increases, the birth rate per 1,000 women aged 15-19 years old 
also increases. This is slightly surprising because we predicted that countries
with high teenage birth rates would have higher female labor participation 
levels. We assumed the young women would be at home, mainly because we saw this
earlier with female educational attainment.  
  
  
**Is the gender develop index related to adolescent birth rate?**

```{r gender-inequality, fig.height=3, fig.width=7, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}
country_joined %>%
  ggplot(aes(x= as.numeric(`Gender Development Index`), 
             y= as.numeric(`Adolescent Birth Rate`))) +
  geom_point(alpha=0.6) +
  labs(
    title= "Is Gender Development Index Related to Adolescent Birth Rate?",
    x="Gender Development Index",
    y="Adolescent Birth Rate"
  )
```

```{r cor-GDI, warning=FALSE, error=FALSE, message=FALSE, echo=FALSE}

cor(as.numeric(country_joined$`Gender Development Index`), 
    as.numeric(country_joined$`Adolescent Birth Rate`), method = "pearson", use="complete.obs")

```

The gender development index measures gender inequality through gender-based 
gaps in education, life expectancy, and incomes, and the closer the value is to 
1, the more equal a society is. We see that as the gender development index 
approaches 1, there is a slight linear decrease (r= -0.5148) in adolescent birth
rate (per 1,000 15-19-year-olds). This relationship makes sense because we 
assume that more equal societies would allow women to make their own 
reproductive choices. They are also more likely to have resources and healthcare
options for women.   
  
  
# Linear regression
  
After looking to see possible relationships, we decided to run a linear 
regression analysis in order to determine which predictors had the most 
influence over teenage pregnancy rates and what that relationship is. The 
general null hypothesis test for every explanatory variable is $\beta\ = 0$, 
which indicates that there is no difference in influence when we control for 
each one individually.  
  
**Which variable has the most significant correlation to adolescent birth rate?**
  
```{r total-linear-regression}
country_joined <- country_joined %>%
  filter(!`GDP` == "..") %>%
  filter(!`Adolescent Birth Rate` == "..") %>%
  filter(!`Maternal Mortality Ratio` == "..") %>%
  filter(!`Mean Years Schooling Female` == "..") %>%
  filter(!`Labour Participation Female` == "..") %>%
  filter(!`Gender Inequality Index` == "..") %>%
  mutate(GDP = as.numeric(`GDP`)) %>%
  mutate(Adolescent_Birth_Rate = as.numeric(`Adolescent Birth Rate`)) %>%
  mutate(Country_new = as.factor(`Country`)) %>%
  mutate(`Maternal Mortality Ratio` = str_replace(`Maternal Mortality Ratio`, ",", "")) %>%
  mutate(Maternal_Mortality_Ratio = as.numeric(`Maternal Mortality Ratio`)) %>%
  mutate(Mean_Years_Schooling_Female = as.numeric(`Mean Years Schooling Female`)) %>%
  mutate(Life_Expectancy_Birth = as.numeric(`Life Expectancy Birth`)) %>%
  mutate(HDI_2019 = as.numeric(`HDI 2019`)) %>%
  mutate(Labour_Participation_Female = as.numeric(`Labour Participation Female`)) %>%
  mutate(Gender_Inequality_Index = as.numeric(`Gender Inequality Index`))

linreg_total <- country_joined %>%
  select(Country_new, Adolescent_Birth_Rate, GDP, Maternal_Mortality_Ratio, Mean_Years_Schooling_Female, Life_Expectancy_Birth, HDI_2019, Labour_Participation_Female, Gender_Inequality_Index)
         
mainfit <- linear_reg() %>%
  set_engine("lm") %>%
  fit(Adolescent_Birth_Rate ~ GDP + Maternal_Mortality_Ratio + Mean_Years_Schooling_Female + Life_Expectancy_Birth + HDI_2019 + Labour_Participation_Female + Gender_Inequality_Index , data = linreg_total)
tidy(mainfit, conf.int=TRUE)

```
  
  
## Interpretation
  
Out of the seven variables we fitted in the linear regression, we see that only 
two explanatory variables (female labor participation and gender inequality 
index) have p-values smaller than our alpha 0.05 (2.395e^-05^ and 8.747e^-07^, 
respectively). Thus, these two variables reject our null hypothesis that there 
is no difference in influence over all of the explanatory predictors.  

To break this down, in a hypothesis test for female labor participation where 
$H0: \beta_6\ = 0$ and $Ha: \beta_6\ != 0$, our rejection of the null hypothesis
indicates an estimated 134.9 increase in adolescent birth rate, holding all 
other variables constant. Additionally, the 95% confidence interval 
($3.332e^-01$, 0.8806) doesn't contain 0.  

In a separate hypothesis test for gender inequality index where 
$H0: \beta_7\ = 0$ and $Ha: \beta_7\ != 0$, our rejection of the null hypothesis
indicates an estimated 0.606 increase in adolescent birth rate, holding all 
other variables constant. Additionally, the 95% confidence interval 
($8.33e^1$, 186.5) doesn't contain 0.  

Furthermore, considering comparatively higher correlation between gender 
inequality index and adolescent birth rate, we decided to further analyze this 
variable in a more detailed scatter plot below.  
  
# Looking further at Gender Inequality Index

```{r HDI Grouping, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
cat_new <- country_joined %>%
  mutate(HDI_final = as.numeric(`HDI Rank.x`)) %>%
  mutate(HDI_cat = cut(
    HDI_final,
    breaks = c(-Inf, 66, 119, 156, 189, Inf),
    labels = c(
      "Very High", "High", "Medium", "Low", "Other")))
```

```{r GII-HDI-birth, echo=FALSE, warning=FALSE, error=FALSE, message=FALSE, fig.width=7, fig.height=3}
ggplot(data = cat_new) +
  geom_point(aes(x = as.numeric(`Gender Inequality Index`), y = as.numeric(`Adolescent Birth Rate`), color = HDI_cat)) +
  labs(title = "Gender Inequality Index vs Adolescent Birth Rate",
       subtitle = "per 1,000 Live Births",
         x = "Gender Inequality Index", 
       y = "Adolescent Birth Rate",
       color = "HDI Development Category")
```

The gender equality index measures the gender disparity in a country on a scale 
from 0 to 1. The greater the value, the greater the measure of inequality.

HDI reflects economic prosperity in a country, so we broke the HDI rankings up 
into four categories to visualize the hypothesis that wealth plays a major roll 
in access to equality. Since gender inequality is such a broad issue we wanted 
to look at wealth desperation in conjunction. This plot supports the idea that 
wealthier countries tend to have lower gender equality disparities and lower 
rates of teen births, and vice versa. Our plot, organized by HDI rank, can help 
us select which countries will benefit most from our intervention targeting 
wealth and gender inequality to improve adolescent birth rates.  

# Summary & Next Steps  

In summation, although we recognize that correlation does not mean causation, 
we recommend that the intervention created to reduce the number of adolescent 
births focus on the communities with the highest rates, countries with low HDI 
rank, and do so by addressing the gender inequality in these countries. Sine 
we have a finite amount of resources, it is best to put them towards low HDI 
countries. Since our highest correlating variable was gender inequality, the 
most efficient way to help adolescent birth rates decrease will be to try and 
close this gap. Some ways that can help close the gender equality gap are 
encouraging girls to stay in school and pursue and education. If girls acquire 
skills that make not just their bodies, but their minds valuable to society then
they will be encouraged to pursue careers and higher degrees, thus waiting to 
have children until they are older. Women who give birth later in life don't 
only have less pregnancy related health risks than teen mothers, but they are 
also more economically stable, and therefore able to give their children a 
better quality of life. 


